Gringo 2.0

### 2008-07-07 doc/documentation/appendix.tex
some documentation changes
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@375 3e02f596-3131-0410-8352-e04eeaec37cc

HIDE_LIST = DOT;
HIDE_LIST = NHIDE_LIST;

@@

CONDITIONAL_LIST = CONDITIONAL_LIST, DDOT, CONDITIONAL;
CONDITIONAL_LIST = DOT;

@@

WEIGHT_LIST = NWEIGHT_LIST;
WEIGHT_LIST = DOT;
NWEIGHT_LIST = NWEIGHT_LIST, COMMA, WEIGHT_TERM;
NWEIGHT_LIST = WEIGHT_TERM;

@@

CONSTR_LIST = NCONSTR_LIST;
CONSTR_LIST = DOT;
NCONSTR_LIST = NCONSTR_LIST, COMMA CONSTR_TERM;
NCONSTR_LIST = CONSTR_TERM;

>>>>

HIDE_LIST = NHIDE_LIST;

@@

CONDITIONAL_LIST = CONDITIONAL_LIST, DDOT, CONDITIONAL;

@@

WEIGHT_LIST = NWEIGHT_LIST;
NWEIGHT_LIST = NWEIGHT_LIST, COMMA, WEIGHT_TERM;
NWEIGHT_LIST = WEIGHT_TERM;

@@

CONSTR_LIST = NCONSTR_LIST;
NCONSTR_LIST = NCONSTR_LIST, COMMA CONSTR_TERM;
NCONSTR_LIST = CONSTR_TERM;

### 2008-07-07 doc/documentation/paper.tex
some documentation changes
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@375 3e02f596-3131-0410-8352-e04eeaec37cc

\documentclass[a4paper,10pt,titlepage]{article}

@@

\paragraph{--convert}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{-l}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{-c}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``--'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output al hold as clasps internal interface currently reads smodels format.
\paragraph{-p}
Parameter ``p'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{-g}
This parameter invokes the new output format for intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. Currently only the ``Conglomeration'' normal form is supported.
\paragraph{default parameters}
The default parameter is ``-l'' for smodels output.
\section{Language Constructs}
The basic elements of a logic program are atoms, constants, variables and rules.
\paragraph{Constant}
A constant is a representation of an entity in the problem domain.
A constant is either a name, starting with a lower case letter or an underscore, or a positive integer.
It also can be encompassed by double quotation marks. When using quotation marks, which is sometimes helpful when using constants with spaces, it has to be remarked that a constant with quotation marks is never equal to a constant without, so ``$a$'' is not $a$.
\paragraph{Variable}
A variable is programming construct to simplify writing logic programs.
It always starts with a capitalized letter and can contain digits and letters.
In the grounding process a variable is replaced by one or more constants.
Each variable occurring in a rule must be domain restricted as described in section \ref{domainrestricted}.
\paragraph{Atom}

@@

\paragraph{Errors}
Gringo2.0 stops the execution and prints an error mesaage to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of atoms that are not omega restricted.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted varibles: { ... }"
\end{verbatim}
This error message gives a list of atoms that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format.
\end{itemize}

@@

\paragraph{Arithmetic Expressions}
The build in arithmetics can be used for integer arithmetic.
As operators addition (+), subtraction (-), multiplication (*), division (/), modulo (mod) and the absolute value operator (abs) can be used.
Furthermore the bitwise operators and(\&), or(?), xor(\textasciicircum) and invert($\sim$) are supported.
Arithmetic expressions can occur everywhere in head and in the body, even in complex constructs.
As parameters, integer constants or expressions that evaluate to integers are used.
The simple example of using operators and range terms
\begin{verbatim}
 a(X-1 .. X+1) :- b(X,Y), c(abs(Y * 2 * -X)).
 b(1,2).
 c(4).
\end{verbatim}
would result in the grounded program\footnote{gringo would simplify the grounded rule to fact, since all literals in the body are facts}
\begin{verbatim}
 a(0) :- b(1,2), c(4).
 a(1) :- b(1,2), c(4).
 a(2) :- b(1,2), c(4).
 b(1,2).
 c(4).
\end{verbatim}

TODO: Bitwise operators !!!!

@@

\section{Benchmarks}
TODO: run on what machine, give reference to encodings.

>>>>

\documentclass[a4paper,10pt]{article}

@@

\paragraph{--convert}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{-l}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{-c}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``--'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output al hold as clasps internal interface currently reads smodels format.
\paragraph{-p}
Parameter ``p'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{-g X}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The X can be any number from $1$ to $7$ to refer to one of the following normalforms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normalform is specified, conglomeration(7)  is assumed.

\paragraph{default parameters}
The default parameter is ``-l'' for smodels output.
\section{Language Constructs}
The basic elements of a logic program are atoms, constants, variables and rules.
\paragraph{Constant}
A constant is a representation of an entity in the problem domain.
A constant is either a name, starting with a lower case letter or an underscore, or a positive integer.
It also can be encompassed by double quotation marks. When using quotation marks, which is sometimes helpful when using constants with spaces, it has to be remarked that a constant with quotation marks is never equal to a constant without, so ``$a$'' is not $a$.
\paragraph{Variable}
A variable is programming construct to simplify writing logic programs.
It always starts with a capitalized letter and can contain digits and letters.
In the grounding process a variable is replaced by one or more constants.
Each variable occurring in a rule must be domain restricted as described in section \ref{domainrestricted}.
\paragraph{Atom}

@@

\paragraph{Errors}
Gringo2.0 stops the execution and prints an error mesaage to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of atoms that are not omega restricted.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted varibles: { ... }"
\end{verbatim}
This error message gives a list of atoms that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format.
\begin{verbatim}
 "Disjunction not allowed on this normalform."
 "Count aggregate/cardinality constraint not allowed
  in this normalform."
 "Count aggregate/cardinality constraint not allowed 
  with non trivial upper bound in this normalform."
 "Count aggregate/cardinality constraint not allowed 
  with negative lower bound in this normalform."
 "Sum aggregate/weight constraint not allowed
  in this normalform."
 "Sum aggregate/weight constraint not allowed 
  with non trivial upper bound in this normalform."
 "Sum aggregate/weight constraint not allowed 
  with negative lower bound in this normalform."
 "Sum aggregate/weight constraint not allowed
  with negative weights in this normalform."
 "Sum aggregate/weight constraint not allowed
  with negative weights in this normalform."
 "Max aggregate not allowed in this normalform."
 "Max aggregate not allowed
  with negative weights in this normalform."
 "Min aggregate not allowed in this normalform."
 "Min aggregate not allowed 
  with negative weights in this normalform."
 "Optimize statement not allowed in this normalform."
\end{verbatim}
%\begin{verbatim}
% "Times aggregate not allowed in this normalform."
% "Times aggregate not allowed
%  with negative weights in this normalform."
%\end{verbatim}
These errors belong to the ASPils output format. Another normal form as to be chosen.
\end{itemize}

@@

\paragraph{Arithmetic Expressions}
The build in arithmetics can be used for integer arithmetic.
As operators addition (+), subtraction (-), multiplication (*), division (/), modulo (mod) and the absolute value operator (abs) can be used.
Furthermore the bitwise operators and(\&), or(?), xor(\textasciicircum) and invert($\sim$) are supported. The bitwise operators are executed on signed integers using the binary complement for negative numbers.
Arithmetic expressions can occur everywhere in head and in the body, even in complex constructs.
As parameters, integer constants or expressions that evaluate to integers are used.
The simple example of using operators and range terms
\begin{verbatim}
 a(X-1 .. X+1) :- b(X,Y), c(abs(Y * 2 * -X)).
 b(1,2).
 c(4).
\end{verbatim}
would result in the grounded program\footnote{gringo would simplify the grounded rules to facts, since all literals in the body are facts}
\begin{verbatim}
 a(0) :- b(1,2), c(4).
 a(1) :- b(1,2), c(4).
 a(2) :- b(1,2), c(4).
 b(1,2).
 c(4).
\end{verbatim}

@@

\section{Benchmarks}
The benchmarks were run on an Athlon XP 1700+ with 1GB of RAM.
TODO: Official (non internal wiki) link to the instances.
